Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-10-04 Latex_files/Thesis/chapters/chapter 3/main.tex
updates 04-10-2024

For a simple circular path one can estimate the sphere's speed by the 
radius of its path and the frequency of its orbit $U = R\omega$; however 
for a more complex path, such as an elliptical orbit the curve needs 
to be parametrised. One can describe the position parameter of a circular 
path as such:

\begin{align}
	r(u) = \left[rcos(2\pi u), rsin(2\pi u), 0 \right]
\end{align}

If we say that $u$ describes time from some initial point we can say $u=t\omega$ 
where omega is the frequency of orbit. Substituting this in and then taking 
the partial derivative of position gives:

\begin{align}
	v(t) = \frac{dr(t)}{dt} = \left[-2\pi r\omega \ sin(2\pi t\omega),
	\ 2\pi r\omega \ cos(2\pi t\omega), 0 \right]
\end{align}

In order to compute U we simply take the magnitude of our velocity. 
For low velocities the fluid flow at the sphere's surface can be computed 
based on its velocity.

\begin{align}
	u_r(r)=-|v(t)|^2\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
\end{align}

@@

Moving a silica bead along an circular path can generate significant fluid 
flow around a larger volume compared to comparable micro-rotors. Using \eqref{eq:galvo_shear} we estimated the shear rate that the surrounding fluid 
would experience at varying speeds. The maximum speed of $7000\ \mu m s^{-1}$
equates to moving the silica bead around a circular path with a frequency of 
$100\ Hz$.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{galvano_shear_rate.png}
	\caption{Shear rate generated by a silica microsphere ($a = 1.57\ \mu m$) 
		in bulk water moving at different speeds. The shear rate is calculated 
		using \eqref{eq:galvo_shear}, with the assumption that the bead is moving
		in a circular path and so the speed is constant through out its path.}
	\label{fig:galvano_shear}
\end{figure}

From figure~\ref{fig:galvano_shear} it is clear that not only is a galvano 
mirror a better option for generating high shear rates but also over a larger 
volume. The only limitation to this method is the effective drag experienced 
by the trapped particle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nucleation with a Stationary and Moving Beam}
As mentioned previously, shearing via optical rotation and particle displacement
did not result in any localised nucleation events even while in the proximity of 
the droplet edge. During the experiments with the galvano-mirror, it was found 
that when no particle was present in the optical trap nucleation events would 
occur while the beam was close to the edge of the droplet, even though the solution
was unsaturated. This has been reported prior \cite{Rungsimanon2010, Liao2022}, 
but was more interesting is how the beam's motion influenced the growth of the 
nucleus.

\subsection{Stationary beam}
\label{sec:stationary}
Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
event in supersaturated glycine solution ($S=1.03$), the beam is a stationary being 
$\approx3.5 \mu m$ from the droplet edge. After a period of roughly 5 minutes a 
nucleus forms at the trap focus, growing quickly (growth rate was approximated using
imageJ to be on the order of $700\ \mu m^2/min$) from the focal point of the trap 
until after roughly $6$ seconds the crystal escapes. Comparing to previous literature
using optical tweezers shows that the growth rate is only loosely connected to the 
solutions supersaturation. Local conditions play a much larger role in the growth 
rate than just the concentration \cite{Flannigan2023}. A likely reason that the trap 
is escaped is due to the fact that crystal is far too large to be held in place and 
is in fact still growing as the solution is supersaturated. The key take away to 
remember is that the beam has no influence over the crystal shape, instead it 
grows outward from the trap focus. Furthermore, due to the fact that the solution
is supersaturated the crystal growth cannot be contained to the trap focus. Instead 
the crystal escapes as its size exceeds the trap focus. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.63\linewidth]{frames_no_beam_movement.pdf}
	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape 
	the trap.}
	\label{fig:stationary_beam}
\end{figure}

\subsection{Moving Beam}
\label{sec:moving}
To test if a rapidly moving silica bead could generate the necessary shear rate 
for crystal nucleation we wanted to see if a trapped silica bead could be trapped
and moved in an aqueous solution. $20\ \mu L$ of glycine and water ($S=1.03$) was 
added to $10\ \mu L$ of a dilute water-silica mixture making the solution 
unsaturated ($S\approx0.7$). However, due to the beam's motion we instead 
encountered several unexpected results.

@@

\subsection{Direct trapping of Glycine clusters}
\label{sec:clusters}
One common aspect involving these droplets is the fact when brought to the laser 
focus the droplets would nucleate immediately \cite{Liao2022}. A solution similar 
to Sec~\ref{sec:stationary} was made up, but without any silica droplets. Once 
again the beam was focused close to the droplet edge, this time the galvano mirror
was scanning a circular path (as shown in figure~\ref{fig:cluster_trapping}). 
After a few minutes of irradiation droplets were seen entering the camera frame, 
because no silica had been added these droplets had to be from the glycine solution. Trapping individual droplets did not result in immediate nucleation even after 
several minutes being trapped. Trying to bring two droplets together resulted in nucleation between the two droplets, compared to \ref{sec:stationary} \& 
\ref{sec:moving} the growth is much slower, taking nearly 40 seconds before the 
crystal structure becomes clearer.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.7\linewidth]{cluster_trapping.pdf}
	\caption{Frames from a longer video demonstrating the trapping of a glycine 
		droplet. Solution is undersaturated glycine and water (S = 0.86), with the 
		laser power is set at $750\ mW$. (a) shows a trapped droplet (outlined in 
		black) being brought into contact with a larger droplet. (b) upon contact 
		a nucleus can be seen between the two droplets. The growth is rather slow 
		with the crystal having no clear defined morphology through (c) and (d). 
		Between frames (e) and (f) the larger droplet finally joins the main crystal.}
	\label{fig:cluster_trapping}
\end{figure}

The fact that these droplets can be trapped indicates they must have a higher 
refractive index than the surrounding solution. It has been shown that the 
concentration of glycine solution is correlated with the refractive indices 
of the liquid \cite{Gowayed2021, Orttung1963}. This suggests that these droplets 
serve to provide solute material to the bulk crystal.

\subsection{Influence of a moving beam front on seed crystals}
One potential application of this phenomena would be in using a 
moving the moving beam front as a method for shaping the final
crystal morphology. Glycine seed crystals were grown via 
evaporative crystallisation over night. The resultant crystals
were as wide as $0.5 mm$ in some cases. Individual crystals were 
collected and suspended in a water + glycine solution ($S = 1.001$)
and the laser was switched on.

@@

>>>>

For a simple circular path one can estimate the sphere's speed by the 
radius of its path and the frequency of its orbit $U = R\omega$; however 
for a more complex path, such as an elliptical orbit the curve needs 
to be parametrised. One can describe the position parameter of a circular 
path as such:
\begin{align}
	r(u) = \left[rcos(2\pi u), rsin(2\pi u), 0 \right]
\end{align}

If we say that $u$ describes time from some initial point we can say $u=t\omega$ 
where omega is the frequency of orbit. Substituting this in and then taking 
the partial derivative of position gives:
\begin{align}
	v(t) = \frac{dr(t)}{dt} = \left[-2\pi r\omega \ sin(2\pi t\omega),
	\ 2\pi r\omega \ cos(2\pi t\omega), 0 \right]
\end{align}

In order to compute U we simply take the magnitude of our velocity. 
For low velocities the fluid flow at the sphere's surface can be computed 
based on its velocity.
\begin{align}
	u_r(r)=-|v(t)|^2\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
\end{align}

@@

Moving a silica bead along an circular path can generate significant fluid 
flow around a larger volume compared to comparable micro-rotors. Using \eqref{eq:galvo_shear} we estimated the shear rate that the surrounding fluid 
would experience at varying speeds. The maximum speed of $7000\ \mu m s^{-1}$
equates to moving the silica bead around a circular path with a frequency of 
$100\ Hz$.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.85\linewidth]{galvano_shear_rate.png}
	\caption{Shear rate generated by a silica microsphere ($a = 1.57\ \mu m$) 
		in bulk water moving at different speeds. The shear rate is calculated 
		using \eqref{eq:galvo_shear}, with the assumption that the bead is moving
		in a circular path and so the speed is constant through out its path.}
	\label{fig:galvano_shear}
\end{figure}

From figure~\ref{fig:galvano_shear} it is clear that not only is a galvano 
mirror a better option for generating high shear rates but also over a larger 
volume. The only limitation to this method is the effective drag experienced 
by the trapped particle. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nucleation with a Stationary and Moving Beam}
As mentioned previously, shearing via optical rotation and particle displacement
did not result in any localised nucleation events even while in the proximity of 
the droplet edge. During the experiments with the galvano-mirror, it was found 
that when no particle was present in the optical trap nucleation events would 
occur while the beam was close to the edge of the droplet, even though the solution
was unsaturated. This has been reported prior \cite{Rungsimanon2010, Liao2022}, 
but was more interesting is how the beam's motion influenced the growth of the 
nucleus.

\subsection{Stationary beam}
\label{sec:stationary}
Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
event in supersaturated glycine solution ($S=1.03$), the beam is a stationary being 
$\approx3.5 \mu m$ from the droplet edge. After a period of roughly 5 minutes a 
nucleus forms at the trap focus, growing quickly (growth rate was approximated using
imageJ to be on the order of $700\ \mu m^2/min$) from the focal point of the trap 
until after roughly $6$ seconds the crystal escapes. Comparing to previous literature
using optical tweezers shows that the growth rate is only loosely connected to the 
solutions supersaturation. Local conditions play a much larger role in the growth 
rate than just the concentration \cite{Flannigan2023}. A likely reason that the trap 
is escaped is due to the fact that crystal is far too large to be held in place and 
is in fact still growing as the solution is supersaturated. The key take away to 
remember is that the beam has no influence over the crystal shape, instead it 
grows outward from the trap focus. Furthermore, due to the fact that the solution
is supersaturated the crystal growth cannot be contained to the trap focus. Instead 
the crystal escapes as its size exceeds the trap focus. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.75\linewidth, height=1.1\linewidth]{frames_no_beam_movement.pdf}
	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape 
	the trap.}
	\label{fig:stationary_beam}
\end{figure}

\newpage
\subsection{Moving Beam}
\label{sec:moving}
To test if a rapidly moving silica bead could generate the necessary shear rate 
for crystal nucleation we wanted to see if a trapped silica bead could be trapped
and moved in an aqueous solution. $20\ \mu L$ of glycine and water ($S=1.03$) was 
added to $10\ \mu L$ of a dilute water-silica mixture making the solution 
unsaturated ($S\approx0.7$). However, due to the beam's motion we instead 
encountered several unexpected results.

@@

\subsection{Direct trapping of Glycine clusters}
\label{sec:clusters}
One common aspect involving these droplets is the fact when brought 
to the laser focus the droplets would nucleate immediately 
\cite{Liao2022}. A solution similar to Sec~\ref{sec:stationary} was 
made up, but without any silica droplets. Once again the beam was 
focused close to the droplet edge, this time the galvano mirror was 
scanning a circular path (as shown in figure~\ref{fig:cluster_trapping}). 
After a few minutes of irradiation droplets were seen entering the 
camera frame, because no silica had been added these droplets had 
to be from the glycine solution. Trapping individual droplets did 
not result in immediate nucleation even after several minutes being 
trapped. Trying to bring two droplets together resulted in nucleation 
between the two droplets, compared to \ref{sec:stationary} \& 
\ref{sec:moving} the growth is much slower, taking nearly 40 seconds 
before the crystal structure becomes clearer.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\linewidth]{cluster_trapping.pdf}
	\caption{Frames from a longer video demonstrating the trapping of a glycine 
		droplet. Solution is undersaturated glycine and water (S = 0.86), with the 
		laser power is set at $750\ mW$. (a) shows a trapped droplet (outlined in 
		black) being brought into contact with a larger droplet. (b) upon contact 
		a nucleus can be seen between the two droplets. The growth is rather slow 
		with the crystal having no clear defined morphology through (c) and (d). 
		Between frames (e) and (f) the larger droplet finally joins the main crystal.}
	\label{fig:cluster_trapping}
\end{figure}

The fact that these droplets can be trapped indicates they must 
have a higher refractive index than the surrounding solution. 
Previous reports have shown that a focused trap will draw in 
solute material from the bulk \cite{Tsuboi2009, Gowayed2021}. It 
has been shown that the concentration of glycine solution is correlated 
with the refractive indices of the liquid \cite{Gowayed2021, Orttung1963}. 
This suggests that these droplets serve to provide solute material 
to the bulk crystal.

\newpage

\subsection{Influence of a moving beam front on seed crystals}
One potential application of this phenomena would be in using a 
moving the moving beam front as a method for shaping the final
crystal morphology. Glycine seed crystals were grown via 
evaporative crystallisation over night. The resultant crystals
were as wide as $0.5 mm$ in some cases. Individual crystals were 
collected and suspended in a water + glycine solution ($S = 1.001$)
and the laser was scanned along a narrow linear path.

As shown by fig.~\ref{fig:seed_crystals} the laser does not in fact 
promote crystal growth but instead forces the crystal to dissolve 
into the bulk solution. The dissolution rate is not as substantial 
as the growth rate seen in previous sections. It takes over 20 
minutes for the for the crystal surface to dissolve more than a 
few microns. Furthermore we do not report any sightings of droplets
close to the surface of the crystal nor do we see anything enter the 
optical trap.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{seed_crystals.pdf}
	\label{fig:seed_crystals}
	\caption{Frames from a seed crystal being irradiated by a scanning
		laser. The laser path is shown below in red. Initially the 
		crystal edge is nearly perpendicular to the laser path. However,
		over the course of nearly twenty minutes the crystal around the 
		laser has dissolved leaving a concave indent.}
\end{figure}

This has some implications: Firstly, the lack of supposed glycine 
droplets suggests that the presence of an interface is crucial in 
order to promote the formation of these droplets. This is consistent
with previous reports on laser nucleation, where a air-solution or 
solution-glass interface is necessary for the trap focus to have any 
noticeable effect on the solute \cite{Gowayed2021, Liao2022, 
Rungsimanon2010}. Secondly, the fact that the seed crystal dissolved 
in a saturated solution suggests that this could be due to a localised
heating effect. A commonly used heating model for optical tweezers is 
the Peterman model \cite{Peterman2003} which accounts for the lateral 
distance between the optical trap and a local heat sink (such as the 
glass cover slip or microscope objective). Even with the heat sink 
the predicted temperature rise in pure water would be around $5-10 K$.  
The fact that the seed crystal dissolves so slowly could simply be due 
to the fact that the local heating is lessened by the moving beam front.

@@

In conclusion, the experimental results from this chapter provide no 
affirmative evidence of shear induced crystal nucleation. This is in
part due to the limited rotational speed that could be achieved using
vaterite microspheres. Simply adding a micro-rotor to a supersaturated
solution will not yield a significant fluid flow to induce crystal 
nucleation. Furthermore, the region in which the local shear rate is 
significant enough is so small that any nucleation events that could 
occur there would not grow large enough to be stable. This does not 
however, suggest that shear induced nucleation is impossible at a 
micro scale. In fact there have been several developments lately that
allow for the precise rotation of multiple rotors to control the 
precise motion of silica beads. The challenge lies in both building the
optical equipment to form multiple strong traps (~ 3 per rotor). These 
traps would have to be strong enough to not only hold the rotors but 
also move them at a sufficiently fast rotational speed that they could
induce a significant enough fluid flow.

On the other hand, the results utilising the galvano mirror have some 
interesting implications for future research into crystal nucleation
using optical tweezers. We noted that in undersaturated conditions the
presence of 'droplets' that could be trapped and later nucleated when 
brought into contact with one another. The fact that these droplets 
could be trapped suggests they have a higher refractive index to the 
solution, given that the solutions were filtered prior to irradiation
dismisses that they could be dust particles. Prior literature suggests
that optical trapping of a supersaturated solution can lead to the 
production of nano droplets \cite{Gowayed2021, Tsuboi2009}. Controlling
the localised growth requires that the growth and trapping are handled
by two separate lasers. Optical traps for amorphous solids has been 
shown as a possibility but the algorithm would need to be adaptive 
to the changing crystal shape.

### 2024-10-04 Latex_files/Thesis/chapters/chapter 4/main.tex
updates 04-10-2024

\section{Characterisation of asymmetric dimers via PSD analysis}
As discussed in \ref{sec:simulated_QPD}, one of the methods developed to 
work in conjunction with \cite{Vigilante2020} is a simulated quadrant photo 
diode for as a position detection system. While it is possible to extract 
all of the relevant dynamical information from a simulation, confirming 
the same behaviour in an experimental setting can be challenging if dealing
with a non-birefringent particle.

As a benchmark we start by considering a single sphere within an optical trap. 
A single polystyrene sphere suspended in water ($a=1\mu m$, $n_p=1.59$, $n_m=1.33$)
is trapped by a focus Gaussian beam and its trajectory was recorded. A 3 second 
is trajectory is a typical measurement time for collecting a power spectrum, 
the spectra was fitted to eq.~\ref{eq:alaised_lorentzian}.
\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{PSD_sphere.png}
	\caption{Recorded power spectra fitted to eq.~\ref{eq:alaised_lorentzian},
		 scattered points represents the blocked data ($n_b=100$). Corner 
		 frequency for the Lorentzian curves are reported in the legend.}
	\label{fig:psd_sphere}
\end{figure}

As shown in fig.~\ref{fig:psd_sphere}, the two power spectra report 
different corner frequencies which would be indicate that the trap 
is not perfectly circular. Using \textit{ott} we can compare the 
expected trap strength to what is reported by a quadrant photo diode:

\begin{center}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting}\\
		\hline
		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
		& \parbox{1.25cm}{\centering 523} & 513 \\
		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{8.16 \%} &
		\multicolumn{2}{|c|}{27.17 \%} &
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline
	\end{tabular}
\end{center}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Characterisation of asymmetric dimers via PSD analysis}
As discussed in \ref{sec:simulated_QPD}, one of the methods 
developed to work in conjunction with \cite{Vigilante2020} 
is a simulated quadrant photo diode for as a position 
detection system. While it is possible to extract all of 
the relevant dynamical information from a simulation, 
confirming the same behaviour in an experimental setting can 
be challenging if dealing with a non-birefringent anisotropic 
scatterer. The QPD is composed of 4 photodiodes that measure
the intensity of light incident on their surfaces. Using 
\textit{ott} we can define a region in the far-field that is 
analogues to the surface of the QPD surface. Evaluating the 
electric field across this surface gives us an approximation
of the QPD signal outputted in an experimental situation. This
does not provide a one-to-one result however as hardware errors
(i.e. internal resistance, external light sources, and vibrations)
would distort the signal somewhat. We can use this as a means 
to examine the limitations of using back-focal plane interferometry
as a means for characterising anisotropic scatterers.

As a benchmark we start by considering a single sphere within an 
optical trap. A single polystyrene sphere suspended in water 
($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focus 
Gaussian beam and its trajectory was recorded. A 3 second is 
trajectory is a typical measurement time for collecting a power 
spectrum, the spectra was fitted to eq.~\ref{eq:alaised_lorentzian}.
\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{PSD_sphere.png}
	\caption{Recorded power spectra fitted to eq.~\ref{eq:alaised_lorentzian},
		 scattered points represents the blocked data ($n_b=100$). Corner 
		 frequency for the Lorentzian curves are reported in the legend.}
	\label{fig:psd_sphere}
\end{figure}

As shown in fig.~\ref{fig:psd_sphere}, the two power spectra report 
different corner frequencies which would be indicate that the trap 
is not perfectly circular. Using \textit{ott} we can compare the 
expected trap strength to what is reported by a quadrant photo diode:
\begin{center}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting}\\
		\hline
		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
		& \parbox{1.25cm}{\centering 523} & 513 \\
		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{8.16 \%} &
		\multicolumn{2}{|c|}{27.17 \%} &
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline
	\end{tabular}
\end{center}
